---
title: "R Notebook"
output: html_notebook
---

# 1. Aim: 
> Analyse complete dataset of results for RA2 lowland for publication

# 2. Data cheking

```{r message = FALSE, echo=FALSE, include=FALSE}
library(data.table)
library (dplyr)
library (ggplot2)
library(RColorBrewer)
library(tidyr)
library(knitr)
```

## Load raw data

```{r ReadRaw}

dataDir <- "C:\\GitHubRepos\\CCII_ShinyApp\\data\\"

rawData <- "Data_RA2_spinUp.csv"

df_raw <- read.csv(file = paste0(dataDir,rawData))     

summary(df_raw)
```

## Tidy up

- Subset data used in the paper only
- RCP8.5 only
- High WHC soils only
- End-century only

```{r}

dfTemp1 <- df_raw %>%
  filter(thisScenario != "midCentury") %>%
  filter(thisRCP != "RCP45") %>%
  filter(thisRCP != "NonAdRCP45") %>%
  filter(thisSoil == "highWHC") %>%
  dplyr::select(-X,-X.1, -skip) %>% # get rid of extra cols
  mutate(Adaptation = factor(ifelse(grepl("NonAd",thisRCP), "Not adapted sow", "Adapted sow"))) %>%
  mutate(Adaptation = factor(Adaptation, levels=c("Not adapted sow","Adapted sow"))) %>%
  mutate(thisRCP = ifelse(!grepl("NonAd",thisRCP),as.character(thisRCP),
                          substring(as.character(thisRCP),6) # get rid of the prefix in RCP label
                          )) %>%
  mutate(thisRCP = factor(thisRCP)) %>%
  mutate(thisScenario = factor(thisScenario, levels=c("base", "endCentury"), 
                               labels=c("Baseline", "End-century"))) %>%
  mutate(TotalBiomass=TotalBiomass/1000) # kg/ha to t/ha


# merge elevation data
altFile <- "C:\\apsim_dev\\Projects\\CCII\\GIS_layers\\DEM_vcsn\\Altitude_Kaituna_RA2.txt"

alt_info <- read.table(altFile, header=TRUE)

dfTemp2 <- merge(dfTemp1, alt_info, by = c("thisLat", "thisLong"))

# merge land use class data FIXME: this is not used early - delete and use late only?
luFile <- "C:\\apsim_dev\\Projects\\CCII\\filter\\LandUse_NZ_All.txt"
lu_info <- read.table(luFile, header=TRUE)
lu_info <- as.data.frame(lu_info)
lu_info$luCode <- as.numeric(lu_info$luCode)
lu_info$fileNo <- NULL

lu_info <- lu_info %>%
  mutate(lu_class = ifelse(luCode <3, "Arable", 
                           ifelse(luCode %in% 3:4, "Moderate",
                           "Non-arable"))) %>%
mutate(lu_class = as.factor(lu_class))

summary(lu_info)

df_worked <- merge(dfTemp2, lu_info, by = c("row", "col"))

# Clean temporary data from memory
df_raw<-dfTemp1 <- dfTemp2 <- lu_info <- alt_info <- NULL

summary(df_worked)

```

## Check time-slices of each scenario

- Shows actual period of analysed data 
- Already excludes spin-up period

```{r}
df_worked %>%
  dplyr::select(thisScenario,thisRCP, year) %>%
  group_by(thisScenario,thisRCP) %>%
  summarise_each(funs(min, max))%>%
  mutate(DataPeriod = max-min)
```

## Data for mapping baseline yields

- Short and long-cycle hybrids

```{r}

outFolder <- "\\\\Lindfs12\\home$\\Cfleit\\My Documents\\My papers\\2016_CCII_Assessment\\maps\\"

mapData_01 <- df_worked %>%
  filter(thisRCP=="ERA") %>%
  group_by(thisLat, thisLong, thisCrop, thisCultivar) %>%
  dplyr::select(thisLat, thisLong, thisCrop, thisCultivar, TotalBiomass) %>%
  mutate(TotalBiomass=TotalBiomass) %>%
  summarise_each(funs(mean))

summary(mapData_01)

write.table(mapData_01, paste0(outFolder,"Map_01_BaselineYields.txt"),quote = FALSE, row.names = FALSE)
  
```

## Get average baseline yields

```{r}

mapData_01 %>%
  group_by(thisCrop,thisCultivar) %>%
  dplyr::select(thisCrop,thisCultivar, TotalBiomass) %>%
  summarise_each(funs(min, max, mean))

```

## How much a short cycle maize affected yields?

```{r}

mapData_01 %>%
  group_by(thisCrop,thisCultivar) %>%
  dplyr::select(thisCrop,thisCultivar, TotalBiomass) %>%
  summarise_each(funs(max)) %>%
  tidyr::spread(thisCultivar,TotalBiomass) %>%
  mutate(shortEffect=(short-long)/long*100)

```

# Calculate climate change impacts on yield 

- Average of all 20 years per grid-cell

```{r}

# isolate GCM BASELINE from RCPs
df_rcp_base <-  df_worked %>%
  filter(thisRCP == "RCPpast") %>%
  mutate(thisCultivar = factor(thisCultivar, levels=(c("short", "long")))) %>%
  mutate(TotalBiomass_Base = TotalBiomass) %>%
   group_by(row, col, thisGCM, thisCrop, thisCultivar) %>%
   dplyr::select(row, col, thisGCM, thisCrop, thisCultivar, TotalBiomass_Base) %>% # name it differently
   summarise_each(funs(mean))
#summary(df_rcp_base)

# isolate GCM FUTURE from RCPs (includes extra fields that we'll use later)
df_rcp_fut <-  df_worked %>%
  filter(thisRCP != "ERA", thisRCP != "RCPpast") %>% # only future scenarios
  mutate(thisCultivar = factor(thisCultivar, levels=(c("short", "long")))) %>%
  mutate(TotalBiomass = TotalBiomass) %>%
   group_by(thisLat, thisLong, row, col, thisScenario, thisGCM, 
                thisCrop, thisCultivar, thisRCP, Adaptation, lu_class) %>%
  dplyr::select(thisLat, thisLong, row, col, thisScenario, thisGCM, 
                thisCrop, thisCultivar, thisRCP, Adaptation, lu_class, Altitude_masl,luCode, TotalBiomass) %>%
  summarise_each(funs(mean))
#summary(df_rcp_fut)

# create df with "only" 20-year average RCP data per pixel now
df_rcp_only <- merge(df_rcp_fut,
                  df_rcp_base,
                  by=c("row", "col", "thisGCM", "thisCrop", "thisCultivar"))
#summary(df_rcp_only)


# calculates climate change impact on biomass
df_bio_diff <- df_rcp_only %>%
  ungroup() %>%
  mutate(diff_CC_abs = TotalBiomass - TotalBiomass_Base) %>%
  mutate(diff_CC_rel = diff_CC_abs/TotalBiomass_Base*100)
 # mutate(Adaptation = factor(Adaptation, levels=c("Not adapted sow", "Adapted sow"))) %>%
 # mutate(thisCultivar = factor(thisCultivar, levels=c("long", "short")))

summary(df_bio_diff)

# Clean intermediary df from memory
df_rcp_base <- df_rcp_fut <- df_rcp_only <- NULL
```

# Map of climate change effect on yields

- Absolute values and percentage of baseline
- Averaged also by the 6 GCMs

```{r}

# Absolute only
mapData_02 <- df_bio_diff %>%
  mutate(Adaptation = factor(gsub(" ","_",Adaptation))) %>%
  dplyr::select(thisLat, thisLong, thisCrop, thisCultivar, Adaptation, diff_CC_abs) %>%
  group_by(thisLat, thisLong, thisCrop, thisCultivar, Adaptation) %>%
  summarise_each(funs(mean))

summary(mapData_02)

write.table(mapData_02, paste0(outFolder,"Map_02_YieldChange_Absolute.txt"),quote = FALSE, row.names = FALSE)

# Percentage only
mapData_03 <- df_bio_diff %>%
    mutate(Adaptation = factor(gsub(" ","_",Adaptation))) %>%
  dplyr::select(thisLat, thisLong, thisCrop, thisCultivar, Adaptation, diff_CC_rel) %>%
  group_by(thisLat, thisLong, thisCrop, thisCultivar, Adaptation) %>%
  summarise_each(funs(mean))

summary(mapData_03)

write.table(mapData_03, paste0(outFolder,"Map_03_YieldChange_Percentage.txt"),quote = FALSE, row.names = FALSE)

```

# Graph CC impact on biomass

```{r, fig.height=8, fig.width=10}

dodge <- position_dodge(width = 1)
dodge1 <- position_dodge(width = 0.8)
dodge2 <- position_dodge(width = 0.5)
dodge3 <- position_dodge(width = 0.25)

# graph all
av_y_diff <- df_bio_diff %>%
  dplyr::select(thisLat, thisLong, thisGCM, thisCrop, thisCultivar, Adaptation, Altitude_masl, diff_CC_rel) %>%
  group_by(thisLat, thisLong, thisGCM, thisCrop,  Adaptation, thisCultivar) %>%
  summarise_each(funs(mean)) 


av_y_diff %>%
   mutate(thisCultivar = factor(thisCultivar,levels=c("short", "long"),
                               labels=c("short-cycle maize","long-cycle maize"))) %>%
  ggplot(aes(x=Adaptation,y=diff_CC_rel)) +
  geom_jitter(alpha=0.5, position = dodge2, aes(colour = factor(thisGCM), size=Altitude_masl)) +
 # geom_jitter(alpha=0.5, position = dodge2, aes(colour = factor(thisGCM))) +
  scale_size(range = c(0, 4)) +
  geom_boxplot(alpha=0.2,position = dodge) +
  geom_violin(alpha=0.2,position = dodge) +
  facet_grid(thisCrop~thisCultivar, scales = "free") + 
  geom_hline(aes(yintercept = 0), linetype = 2) +
  ylab(" Climate change impact \non crop yields (% baseline)")+
  xlab("Sowing date adaptation")  +
  theme(legend.position="right") +
  labs(size="Elevation (m)", colour="GCMs") +
 # theme(legend.title=element_blank()) +
  theme(text = element_text(size=18))
```

# get numbers of CC impact

```{r}
av_y_diff %>%
  ungroup() %>%
  group_by(thisCrop,Adaptation,thisCultivar) %>%
  summarise(min=min(diff_CC_rel),
            max=max(diff_CC_rel),
            median=median(diff_CC_rel)
            )
  
```


# Calculate average response to adaptation

- 

```{r}

```




## combination of wheat and maize responses

```{r, echo=FALSE, warning=FALSE, fig.height=6, fig.width=12}

l_s <- 5

# get GCM variability too
av_sd_y_diff <- df_bio_diff %>%
  dplyr::select(thisLat, thisLong, thisCrop, 
                thisCultivar, Adaptation, Altitude_masl, diff_CC_abs) %>%
  group_by(thisLat, thisLong, thisCrop, Adaptation, Altitude_masl, thisCultivar) %>%
  summarise_each(funs(mean, sd))

mx<-max(av_sd_y_diff$mean)
mn<-min(av_sd_y_diff$mean)

# summary(av_sd_y_diff)

# Analyse trade-off between maize and wheat yields
av_sd_y_diff %>%
    mutate(thisCultivar = factor(thisCultivar,levels=c("short", "long"),
                               labels=c("Short-cycle maize","Long-cycle maize"))) %>%
  unite(temp,mean,sd) %>%
  tidyr::spread(thisCrop, temp) %>%
  separate(maize, c("maize_mean","maize_sd"), sep="_") %>%
  separate(wheat, c("wheat_mean","wheat_sd"), sep="_") %>%
  mutate(maize_mean = as.numeric(maize_mean),
         wheat_mean = as.numeric(wheat_mean),
         maize_sd = as.numeric(maize_sd),
         wheat_sd = as.numeric(wheat_sd)) %>%
  ggplot(aes(x=maize_mean, y=wheat_mean, colour=factor(thisCultivar))) +
  annotate("rect", xmin = mn, xmax = 0, ymin = mn, ymax = 0, alpha=0.05, fill="red") +
  annotate("rect", xmin = 0, xmax = mx, ymin = 0, ymax = mx, alpha=0.05, fill="green") +
  annotate("rect", xmin = mn, xmax = 0, ymin = 0, ymax = mx, alpha=0.05, fill="yellow") +
  annotate("rect", xmin = 0, xmax = mx, ymin = mn, ymax = 0, alpha=0.05, fill="yellow") +
  geom_point(aes(size = as.numeric(Altitude_masl), shape=thisCultivar), alpha=1) +
  scale_shape(solid = FALSE) +
  facet_wrap(~Adaptation, scales = "free") + 
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 0, linetype = 2) +
  xlab("Maize silage (t/ha)") + 
  ylab("Wheat cover crop (t/ha)") + 
  geom_errorbarh(aes(xmin=maize_mean-maize_sd, xmax=maize_mean+maize_sd),colour="grey", width=.1, alpha=0.8) +
  geom_errorbar(aes(ymin=wheat_mean-wheat_sd, ymax=wheat_mean+wheat_sd), colour="grey", width=.1, alpha=0.8) +
  labs(size="Elevation (m)", colour="Maize genotype", shape="") +
  guides(shape=FALSE) +
  xlim(mn,mx) +
  ylim(mn,mx) + 
  theme_bw() +
  theme(text=element_text(size=18))  +
  ggtitle("Climate change impact \non above-ground yield") +
  theme(plot.title = element_text(hjust = 0.5))

```


## Do total crop biomass per rotation

- sum up absolute climate change impact of maize and wheat

```{r}

# set df
df_bio_diff_tot <- av_sd_y_diff %>%
  unite(temp,mean,sd) %>%
  tidyr::spread(thisCrop, temp) %>%
  separate(maize, c("maize_mean","maize_sd"), sep="_") %>%
  separate(wheat, c("wheat_mean","wheat_sd"), sep="_") %>%   
  mutate(maize_mean = as.numeric(maize_mean),
         wheat_mean = as.numeric(wheat_mean),
         maize_sd = as.numeric(maize_sd),
         wheat_sd = as.numeric(wheat_sd)) %>%
  mutate(totBioDiff = maize_mean + wheat_mean) %>%
  mutate(totSdDiff = maize_sd + wheat_sd)

summary(df_bio_diff_tot)

```

## merge with land use data

```{r}

# merge land use class
df_lu <- df_worked %>% dplyr::select(thisLat, thisLong,lu_class) %>% distinct()
df <- merge(df_bio_diff_tot,df_lu, by=c("thisLat","thisLong"))

summary(df)


```

```{r, fig.height=6, fig.width=12}


# find ranges
mx<-max(df_bio_diff_tot$totBioDiff)
mn<-min(df_bio_diff_tot$totBioDiff)

# graph
df %>%
  mutate(thisCultivar = factor(thisCultivar, levels=c("short", "long"), 
                                labels=c("short-cycle maize", "long-cycle maize"))) %>%
  ggplot(aes(x=Altitude_masl,y=totBioDiff, colour=Adaptation)) +
  geom_smooth(alpha=0.2, aes(linetype=Adaptation, fill=Adaptation)) +
  geom_point(size=3, aes(shape=lu_class)) +
  scale_shape(solid = FALSE) +
  facet_grid(.~thisCultivar) +
  geom_hline(yintercept = 0, linetype = 2) +
  ylab("Climate change impact on \naccumulated crop rotation yield (t/ha)") +
  xlab("Elevation (m above sea level)") +
  theme(text=element_text(size=18))+
  labs(colour="Adaptation", shape="Land use class") +
  geom_errorbar(aes(ymin=totBioDiff-totSdDiff, ymax=totBioDiff+totSdDiff, colour=Adaptation), width=.1, alpha=0.8)
```

## How do relative and absolute yield changes compare?

- used in discussion

```{r}
df_bio_diff %>%
  group_by(thisCrop) %>%
  ggplot(aes(x=TotalBiomass_Base, y=diff_CC_rel, colour=thisCrop)) +
  geom_point()+
  facet_wrap(~thisCrop, scales = 'free')
```

```{r}
df_bio_diff %>%
  group_by(thisCrop) %>%
  ggplot(aes(x=diff_CC_rel, y=diff_CC_abs, colour=thisCrop)) +
  geom_point()+
  facet_grid(.~thisCrop, scales = 'free')
```

## How much catch crop removed

```{r}
df_worked %>%
  filter(thisCrop=="wheat") %>%
  dplyr::select(thisScenario,Adaptation, thisGCM, thisCultivar,thisSoil,PTleach,PTcropn ) %>%
  group_by(thisScenario,Adaptation, thisGCM, thisCultivar, thisSoil) %>%
  summarise_each(funs(mean))  %>%
  summary()
  
```

```{r}
df_worked %>%
  filter(thisCrop == "maize") %>%
    dplyr::select(thisScenario,Adaptation, thisGCM, thisCultivar,thisSoil,HarvestIndex ) %>%
  group_by(thisScenario,Adaptation, thisGCM, thisCultivar, thisSoil) %>%
  summarise_each(funs(mean))  %>%
  tidyr::spread(Adaptation,HarvestIndex) %>%
  summary()
```

## Table with quality and environmental aspects

```{r}
df_worked %>%
  filter(thisCrop == "maize") %>%
  filter(thisGCM !="ERA") %>%
  dplyr::select(thisLat,thisLong,thisScenario,Adaptation, thisCultivar,PTcropn, HarvestIndex) %>%
  group_by(thisLat,thisLong,thisScenario,Adaptation, thisCultivar) %>%
  summarise_each(funs(mean)) %>%
  tidyr::gather(VarName,VarValue, PTcropn:HarvestIndex) %>%
  ggplot(aes(x=Adaptation,y=VarValue)) +
  geom_boxplot() +
  facet_wrap(VarName~thisScenario, ncol=2, scales = 'free') +
  geom_jitter(aes(colour=thisCultivar))
```

```{r, fig.height=6, fig.width=12}

# plot against average temperature in the pixel

df_worked %>%
  filter(thisCrop == "wheat") %>%
  filter(thisGCM !="ERA") %>%
  filter(Adaptation =="Adapted sow") %>%
  dplyr::select(thisLat,thisLong,year,Altitude_masl, thisScenario,Adaptation, thisCultivar,PTcropn) %>%
  group_by(thisLat,thisLong,year,Altitude_masl, thisScenario,Adaptation, thisCultivar) %>%
  summarise_each(funs(mean,sd)) %>%
  ggplot(aes(x=thisLat,y=mean,colour=thisCultivar))+
  geom_point(alpha=0.2, aes(size=Altitude_masl)) +
#  geom_boxplot() +
  geom_smooth(aes(fill=thisCultivar)) +
  facet_grid(Adaptation~thisScenario, scales = 'free') 
```

