---
title: "MetFileGraphs"
output: html_document
---

## Average per pixel

- Requires MileFileSummary.R to be run first
- Requires MileFileReadAll.Rmd to be run first
- sort out units
- graph averages
- Derived from original at:C:\apsim_dev\Projects\CCII\RA2_CaseStudy\source\MetFileGraphs(RA2).Rmd

```{r, warning=FALSE, include =FALSE}
#devtools::install_github("dkahle/ggmap")
#devtools::install_github("hadley/ggplot2")
library(tcltk2)
library(dplyr)
library(data.table)
library(raster)
library(ggplot2)
library(maptools) # for shapefiles
library(ggmap)
```

Read data
```{r}

folderName <- "C:\\apsim_dev\\Projects\\CCII\\RA2_CaseStudy\\metFiles\\"
workDirName <- "C:\\GitHubRepos\\2017_CCII_RA2_paper\\"
fileName <- "All_Met_Data.csv"

DATA <- read.csv(paste0(folderName,fileName))
summary(DATA)

```

# Data clean up

- for publication
- RCP8.5 only
- end-century only


```{r SummaryAll, echo=FALSE,echo=FALSE, fig.height=18,fig.width=12}

# FIXME: this will have to deal with more RCPs in teh future

# average across grid-cells
df <- DATA %>%
  mutate(timeSlice = as.factor(paste0("ts_",timeSlice))) %>% # avoid start of factor name with number
  group_by(row, col, gcmStp, rcpStp, timeSlice) %>% 
  summarise_all(funs(mean)) %>%
  dplyr::select(-X, -X.1,-skip,-fileNo, -year, -day) %>% # must remove otherwise
  mutate(rain = rain*365) %>% # change to annual values , radn = radn*365
  tidyr::gather("Variable", "VarValue", maxt:co2) %>%
  mutate(Variable = as.factor(Variable)) %>%
  mutate(VariableName = factor(Variable))%>%
  mutate(Variable = factor(Variable,levels= c("co2","mint","maxt","rain","radn"),
                           labels=c("co2","Minimum temperature",
                                   "Maximum temperature",
                                   "Rainfall",
                                   "Solar radiation"
                                  ))) %>%
#  mutate(Variable = factor(Variable,levels= c("co2","mint","maxt","rain","radn"))) %>%
  mutate(timeSliceName = factor(timeSlice)) %>%
  mutate(timeSlice = factor(timeSlice,levels = c("ts_a_base","ts_b_midCentury","ts_c_endCentury"), 
                            labels=c("Baseline", "Mid-century", "End-century")))


# subset RCP and time-slice for publication
av.per.pixel <- df %>%
  filter(rcpStp == "RCP85")%>%
  filter(timeSlice != "Mid-century")


summary(av.per.pixel)

#x <- c("PM[10]","SO[2]", "NO", "NO[2]")

```

# Check CO2 values

- atmospheric CO2 average

```{r CO2}
# what's average co2 for each time slice?
av_co2 <- av.per.pixel %>% 
  dplyr::filter(Variable =="co2") %>%
  group_by(rcpStp,timeSlice,Variable) %>% 
  summarise(mean=round(mean(VarValue),0))

print(av_co2)

```

## Check averages

- average of all pixels
- average of all GCMs
- save averages

```{r}
# save for analysis
#write.csv(av.per.pixel, file = paste0(workDirName,"AvsPerPixel_CCII_paper.csv"), row.names = FALSE)

# find averages to plot as ref
avs1 <- av.per.pixel %>% 
  dplyr::filter(Variable !="co2") %>%
  group_by(timeSlice,rcpStp,Variable) %>% 
  summarise(mean=mean(VarValue)) 

avs1 %>%
  tidyr::spread(timeSlice,mean)
```

# Data for map in supplement

```{r}
av_cv_abs <- av.per.pixel %>% 
  dplyr::filter(Variable !="co2") %>%
  group_by(thisLat,thisLong,gcmStp, timeSlice,Variable) %>% 
  summarise(av_pix=mean(VarValue)) %>% # first get averages by GCM
  group_by(thisLat,thisLong,timeSlice,Variable) %>% # now do cv by GCM
  summarise(gcm_mean=mean(av_pix), 
            stDev = sd(av_pix),
            cv=stDev/gcm_mean*100) 

summary(av_cv_abs)

write.csv(av_cv_abs, file = paste0(workDirName,"ClimateDataPerPixel_av_cv_CCII_RA2.csv"), row.names = FALSE)
```



## Graph for supplement

```{r, fig.height=8, fig.width=12}


# graph averages
av.per.pixel  %>%
  dplyr::filter(Variable !="co2") %>%
  ggplot(aes(x=gcmStp,y=VarValue, colour=factor(gcmStp))) +
  geom_boxplot(alpha=0.5) +
  geom_violin(alpha=0) +
  geom_jitter(size=0.5) +
  facet_wrap(timeSlice~Variable, scales = "free",  ncol = 4) + #,labeller = label_parsed
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme(legend.position="none") +
  xlab("Climate model") +
  ylab("Variable value") + 
  geom_hline(aes(yintercept=mean),avs1,linetype = 2)
```

## Time-slice diffs per GCM within grid cell

Create difference of time-slice averages per pixel (valid for GCM only)

```{r DiffPerPixel, echo=FALSE,echo=FALSE, fig.height=18,fig.width=12}

# diff.timeSlice.per.pixel
diff.timeSlice.per.pixel <- av.per.pixel %>%
  dplyr::select(-timeSlice)%>%
  tidyr::spread(timeSliceName, VarValue)  %>%
  # mutate(diff_mid_a = ts_b_midCentury - ts_a_base,
  #        diff_end_a = ts_c_endCentury - ts_a_base,
  #        diff_mid_r = ts_b_midCentury/ts_a_base*100-100,
  #        diff_end_r = ts_c_endCentury/ts_a_base*100-100
  #        ) %>%
    mutate(diff_end_a = ts_c_endCentury - ts_a_base,
         diff_end_r = ts_c_endCentury/ts_a_base*100-100
         ) %>%
  tidyr::gather("Diff_type", "DiffValue", diff_end_a:diff_end_r) %>%
  mutate(Diff_type = as.factor(Diff_type)) %>%
  mutate(Diff_typeName = factor(Diff_type,
                            levels=c("diff_end_a",
                                     "diff_end_r"),
                            labels=c("End-century (absolute)", 
                           "End-century")))

#head(diff.timeSlice.per.pixel)
summary(diff.timeSlice.per.pixel)

write.csv(diff.timeSlice.per.pixel, file = paste0(folderName,"DiffPerPixel.csv"))

```
Subset for map 

- Only absolute differences
- 
```{r}
# Subset of data for Anne-gaelle for Supplement 

gcm.var.per.pixel.subset <- diff.timeSlice.per.pixel %>%
  filter(Diff_type == "diff_end_a") %>%
  filter(VariableName != "co2") %>%
 # filter(rcpStp == "RCP85")%>%
  dplyr::select(-Diff_typeName,-ts_a_base,-ts_c_endCentury)

summary(gcm.var.per.pixel.subset)

write.csv(diff.timeSlice.per.pixel, file = paste0(folderName,"DiffPerPixel_subset.csv"), row.names = FALSE)

```



# Show distribution of baseline + RCP 8.5 changes

- Used in paper
- RCP 8.5
- End-century only

```{r}

# create subset for paper
temp1 <- diff.timeSlice.per.pixel %>%
  ungroup() %>%
  filter(rcpStp == "RCP85") %>% # for paper
  filter(Variable != "co2") %>% 
  filter(Diff_typeName == "End-century (absolute)") %>%
  mutate(gridCell = (paste0(row,"_", col))) %>%
  mutate(gridCell=as.factor(gridCell)) %>%
  mutate(GCMs = gcmStp) %>%
  mutate(BaseValue = ts_a_base) %>%
  dplyr::select(row, col, Variable, GCMs, BaseValue, DiffValue) %>%
  tidyr::gather(VarType, VarValue,BaseValue:DiffValue) %>%
  mutate(VarType = factor(VarType, levels = c("BaseValue","DiffValue"), 
                          labels=c("Baseline value","Climate change impact") ))

str(temp1)

summary(temp1)

```

```{r}
# find averages
avs4 <- temp1 %>% 
  group_by(Variable, VarType) %>% 
  summarise(mean=mean(VarValue))

summary(avs4)




```
# find mode value
```{r}
# Create the model function.
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

modeV <- temp1 %>% 
  group_by(Variable, VarType) %>% 
  summarise(mode=getmode(VarValue))

summary(modeV)
```



```{r, fig.height=10, fig.width=12}

# plot
temp1 %>%
  ggplot(aes(VarValue, fill = GCMs, colour = GCMs)) +
  geom_density(alpha = 0.2, aes(y=..scaled..)) +
  facet_wrap(Variable~VarType, scales = "free", ncol = 2) +
    xlab("Average weather variables for 20-year analysis period") +
  ylab("Distribution across catchment area" ) + 
  geom_vline(aes(xintercept=mean),avs4,linetype = 2) +
  geom_vline(aes(xintercept=mode),modeV,linetype = 3)
```

# Show values for writing materials and methods in paper

```{r}

temp1 %>%
  tidyr::spread(Variable, VarValue) %>%
  filter(VarType == "Baseline value") %>%
  summary()

```

# show differences now

```{r}
temp1 %>%
  tidyr::spread(Variable, VarValue) %>%
  filter(VarType == "Climate change impact") %>%
  summary()
```

## GCM diffs per "time-slice deltas" within grid cell

- it only makes sense to compare the variability in "deltas" as the absolute values are inherently different among GCMs 

```{r rasteriseAverages}

# group all GCMs (gcm variability related to time-slice diffs)
gcm.var.per.pixel <- diff.timeSlice.per.pixel %>%
  filter(Variable != "co2") %>%
 # mutate(gridCell = as.factor(paste0(row,"_", col))) %>%
  group_by(thisLat, thisLong, rcpStp, Variable, Diff_type) %>%
  dplyr::select(thisLat, thisLong, rcpStp, Variable, Diff_type,DiffValue) %>%
  summarise_all(funs(mean,sd)) %>%
  mutate(cv = sd/mean*100) %>%
  tidyr::gather("Stats","StatsValue",mean:cv) %>%
  mutate(Stats = as.factor(Stats))

#head(gcm.var.per.pixel)
summary(gcm.var.per.pixel)

write.csv(gcm.var.per.pixel, file =paste0(folderName,"RasterOutput(RA2)_stats_GCMs.csv"), row.names = FALSE)


```

## Check GCM variability

- Unfinished
- FIXME:ggmap not working anymore
- A new script can read "RasterOutput(RA2)_stats_GCMs.csv" and apply ggmap

```{r, fig.height=30, fig.width=12}
# gcm.var.per.pixel %>%
#   filter(Stats == "cv") %>%
#   ggplot(aes(x=StatsValue,y=thisPixel)) +
#   geom_point() +
#   facet_wrap(Variable~Diff_type+rcpStp, scales="free")
```



## Differences among GCM (regarding relative changes projected for time-slices)

```{r SpatialGcmUncertanty}

# # Not yet understanding ggmap synthax
# 
# 
# 
# 
# # get shape
# library(maptools)
# shpFile <- "C:\\apsim_dev\\Projects\\CCII\\GIS_layers\\CaseStudy\\lowerKaitunabnd(WGS84).shp"
# shpData <- readShapeSpatial(shpFile, proj4string = CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
# str(shpData) 
# 
# # set base map
# library(ggmap)
# lat <- -37.925
# long <- 176.275
# zoom <- 10
# 
# baseMapData <- get_map(location = c(lon = long, lat = lat),
#                          color = "color",
#                          source = "google",
#                          maptype = "roadmap",
#                          zoom = zoom)
# 
# baseMap <- ggmap(baseMapData,
#       extent = "device",
#       ylab = "Latitude",
#       xlab = "Longitude")
# 
# # FIXME: not working yet
# head(gcm.var.per.pixel)
# 
# # rainfall
# mapDf <- gcm.var.per.pixel %>%
#   filter(Diff_type == "diff_end_r" & Variable == "rain") %>%
#   filter(Stats == "mean")
# 
# 
# baseMap +
#   geom_point(aes(x=thisLong, y=thisLat, show_guide=TRUE, colour=StatsValue), size = 8, 
#              data=mapDf, alpha=0.9, na.rm=T, shape = 15) +
#  # scale_color_gradient(low='blue', high='red') +
#   geom_polygon(aes(x=long, y=lat, group=id), data=shpData, colour="black",fill=NA) +
#   facet_grid(~Variable) + 
#     scale_colour_gradient2(low="#deebf7", mid = "#9ecae1", 
#                          high="#3182bd", midpoint=0)






# baseMap +
#   stat_density2d(aes(x=thisLong, y=thisLat, z=StatsValue, fill = ..density..), geom="tile", 
#                  contour = FALSE, data=mapDf) +
#   geom_polygon(aes(x=long, y=lat, group=id), data=shpData, colour="black",fill=NA) +
#   facet_grid(~Variable) + 
#   scale_colour_gradient2(low="#ef8a62", mid = "#ffffbf", 
#                          high="#67a9cf", midpoint=0)


```

```{r}
# # temp
# mapDf <- gcm.var.per.pixel %>%
#   filter(Diff_type == "diff_end_a" & Variable == "maxt") %>%
#  # filter(!grepl("_a", Diff_type)) %>% # get rid of abaolute diffs
#   filter(Stats == "mean")
# 
# 
# baseMap +
#   geom_point(aes(x=thisLong, y=thisLat, show_guide=TRUE, colour=StatsValue), size = 7, 
#              data=mapDf, alpha=0.9, na.rm=T, shape = 15) +
#  # scale_color_gradient(low='blue', high='red') +
#   geom_polygon(aes(x=long, y=lat, group=id), data=shpData, colour="black",fill=NA) +
#   facet_grid(~Variable) + 
#     scale_colour_gradient2(low="#fee6ce", mid = "#fdae6b", 
#                          high="#e6550d", midpoint=median(mapDf$StatsValue)) +
#   guides(fill = guide_legend(keywidth = 0.5, keyheight = 1, title = "oC"))
```

```{r}
# # trying panel
# mapDf <- gcm.var.per.pixel %>%
#   filter(Diff_type == "diff_end_r") %>%
#   filter(Stats == "mean")
# 
# 
# baseMap +
#   geom_point(aes(x=thisLong, y=thisLat, show_guide=TRUE, colour=StatsValue), size = 4, 
#              data=mapDf, alpha=0.9, na.rm=T, shape = 15) +
#   geom_polygon(aes(x=long, y=lat, group=id), data=shpData, colour="black",fill=NA) +
#   facet_grid(~Variable) + 
#     scale_colour_gradient2(low="#fee6ce", mid = "#fdae6b", 
#                          high="#e6550d", midpoint=median(mapDf$StatsValue))
```

## Rasterise diifs - maybe not needed anymore after ggmap works

```{r RasterGcmDiffs}

# head(mapDf) # already selected for relative diffs
# 
# outNames <- as.character(unique(mapDf$Variable))
# scenNames <- as.character(unique(mapDf$Diff_type))
# statsNames <- as.character(unique(mapDf$Stats))
# 
# par(mfcol=c(1,1))
# 
# # creates directories to save results
# for (st in 1:length(statsNames)) {
# #  dir.create(file.path(getwd(),scenNames[sc]), showWarnings = FALSE)
#   dir.create(file.path(getwd(),"MapsDiff"), showWarnings = FALSE)
# }
# 
# mapNo <- 0
# 
# spg <- list() # dataframe list to be coerced as spatial object
# rast <- list() # raster
# 
# for (st in 1:length(statsNames)) {
# 
#  for (sc in 1:length(scenNames)) {
# 
#    for(o in 1:length(outNames)){
#      
#      mapNo <- mapNo + 1
#      
#       df <- NULL # data subset
#       df <- mapDf %>%
#       subset(Diff_type == scenNames[sc] & 
#                Variable == outNames[o] & 
#                Stats == statsNames[st])
#       
#       # Create grid type objects stored by map #
#       spg[[o]] <- data.frame(df$thisLong, df$thisLat, df$StatsValue)
#       coordinates(spg[[o]]) = ~ df.thisLong + df.thisLat # Attention to variable names
#       gridded(spg[[o]]) <- TRUE
#       rast[[o]] <- raster(spg[[o]])
#       proj4string(rast[[o]]) = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
#       
#       #Title
#       tit <- paste0(mapNo,"_",scenNames[sc],"_",outNames[o],"_",statsNames[st]) 
#       
#       # save it as image
#       thisFile <- paste0(getwd(),"//MapsDiff//",tit,".tiff")
#         writeRaster(rast[[o]], file= paste0(thisFile), overwrite=TRUE) 
#       
#      # plot it
#      par(mar=c(3,3,3,3))
#      par(mfrow=c(1,2))
#      plot(rast[[o]])
#      boxplot(rast[[o]],main = tit)
#         
#     
#       # stack them together
#       if(mapNo == 1) {
#         s <- NULL
#         s <- stack(rast[[o]])} 
#      else {
#         s[[mapNo]] <- rast[[o]]
#       }
#      
#      print(tit)
#      
#    } # statsNames
#    
#  } # scenNames
#   
# } # outNames
# 
# 
# # plot maps
# 
# par(mfrow=c(2,2))
# for (mc in 1:mapNo) {
#   plot(s[[mc]], main = outNames[mc])
#  # writeRaster(rast_diff, file=paste0("diffAbs_",outNames[mc],".tiff"), overwrite=TRUE) 
# }
```

